Selective stock enhancement undermines community stability

### 2021-12-27 rmd/manuscript.Rmd
Update manuscript.Rmd

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and excessive release of particular species may have unforeseen consequences for ecological communities by disrupting the sensitive balance of species interactions. Evidence for this hypothesis is lacking, however.

@@

@@

$f_R~(\ge 0)$ is the density-independent survival of released individuals relative to wild individuals (the relative fitness). Therefore, the model accounts for the reduced fitness of captive-bred individuals due to genetic effects and/or phenotypic plasticity while considering the reproductive contribution of released individuals to the next generation of the enhanced species.

@@

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is sufficiently weak to allow for stable coexistence of competing species in the absence of stock enhancement.

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Stable coexistence of competing species rests largely on niche difference in consumable resource use ($\alpha_{ij, i \ne j} < 1$) because it creates a situation where numerically dominant species undergoes stronger intraspecific competition. However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Such excessive individuals impose additional intra- and interspecific competition. Consequently, stock enhancement compromises the ecological process giving rise to "overyielding", in which total community density of an ecologically diverse community exceeds what would be expected in a single species community ($\sum_i^SN_i > K$). In an extreme scenario, intensive stock enhancement extirpates competing species, reducing biodiversity in the recipient community.

We must note that stock enhancement helps increase the density of the enhanced species when the intrinsic population growth rate is significantly small. This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species. Even so, community level response was XXX, limiting the .

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on long-term stability of stream fish communities. Masu salmon is one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of stock enhancement. Therefore, the dataset sets the stage for a "natural experiment" to test the theoretical predictions. In this study, we assembled time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for confounding effects of climates and local abiotic factors.

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released. The positive relationship between the CV and the number of fish released was strongly supported by a hierarchical linear model, in which the probability of the regression coefficient being positive was XXX. This was associated with the reduced long-term averages of population densities of enhanced (masu salmon) and unenhanced fish groups. Both fish groups showed low population densities with high numbers of fish released (STATS), although statistical support for this relationship in the individual fish groups was weak to moderate (STATS). In the meantime, the SDs of population densities had vague relationships with the number of fish released, and neither positive or negative relationship was supported by statistical analysis.

Our study system satisfies key theoretical assumptions that are required for the destabilizing effect of selective stock enhancement to emerge. The first assumption is that released fish compete with wild fish. This type of competition is very likely to occur in the study streams. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that this species compete with other species of stream fish. In particular, masu salmon raised in hatchery environments are more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild. Second, our theory predicts the emergence of destabilizing effect when the number of fish released is externally controlled. This is exactly what was observed in our study system. In the protected watersheds, there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). Instead, these released fish are likely controlled by socioeconomic factors, such as budget allocation to hatchery programs in this region. As such, released fish are probably "excessive" given the density-dependent regulation of natural communities, imposing intense resource competition that would otherwise not exist.

The results of this study point to two important avenues for biodiversity conservation. First, release of captive-bred individuals should be avoided when strong density-dependence operates in the recipient ecosystem. Although captive breeding programs carefully consider the fitness of captive-bred individuals in the wild, it often fails to account for the capacity of recipient ecosystems to support biodiversity. Negative density dependence is a general regulating mechanism of biodiversity, and the ignorance of the ecosystem capacity may deteriorate biodiversity, including the supplemented species. Therefore, considering this aspect is crucial for saving biodiversity from unexpected effects of stock enhancement. Second, we should prioritize the preservation and restoration of ecosystems. Myriads of human influences have caused substantial loss and degradation of natural habitats. The lack of conservation actions that ameliorate these pressures will impose wild individuals to compete with captive-bred individuals for limited resources, further worsening the current ecosystem state. Collectively, without the broader framework that recognizes the regulation process of biodiversity, stock enhancement will never be effective but leads to biodiversity loss.

>>>>

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species underpins emergent stability of ecological communities. In this context, stock enhancement represents excessive individuals that deviate from the internal regulation process of self-sustaining communities; therefore, it may disrupt the sensitive balance of species interactions with unforeseen consequences for community dynamics. Evidence for this hypothesis is lacking, however.

@@

## Theoretical prediction

@@

$f_R~(\ge 0)$ is the density-independent survival of released individuals relative to wild individuals (the relative fitness). Therefore, the model accounts for fitness difference of captive-bred individuals due to genetic effects and/or phenotypic plasticity when considering the reproductive contribution of released individuals to the next generation of the enhanced species.

@@

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is very weak.

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Stable coexistence of competing species rests largely on niche difference in consumable resource use ($\alpha_{ij, i \ne j} < 1$) because it creates a situation where numerically dominant species undergoes stronger intraspecific competition. However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Such excessive individuals impose additional intra- and interspecific competition. Consequently, stock enhancement compromises the ecological process giving rise to "overyielding", in which total community density of an ecologically diverse community exceeds what would be expected in a single species community ($\sum_i^SN_i > K$). In an extreme scenario, intensive stock enhancement extirpated competing species, reducing biodiversity in the recipient community.

We must note that stock enhancement helps increase the density of the enhanced species when the intrinsic population growth rate is significantly small ($r_1 < 0.5$). This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species. Yet, the community-level effect is still destabilizing due to the intensified competition with unenhanced species. Therefore, managers should consider the propagating effect of captive breeding when implementing their programs.

## Empirical evidence

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on long-term stability of stream fish communities. Masu salmon is one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of stock enhancement. Therefore, the dataset sets the stage for a "natural experiment" to test our theoretical predictions. In this study, we used time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for potential confounding factors (climates and local abiotic factors).

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released. The positive relationship between the CV and the number of fish released was strongly supported by a hierarchical linear model, in which the probability of the regression coefficient being positive was XXX. This result was associated with the reduced long-term averages of population densities of enhanced (masu salmon) and unenhanced fish groups. Both fish groups showed low population densities with high numbers of fish released (STATS), although statistical support for this relationship in the individual fish groups was weak to moderate (STATS). In the meantime, the SDs of population densities had vague relationships with the number of fish released, and neither positive or negative relationship was supported by statistical analysis.

Our study system satisfies key theoretical assumptions that are required for the destabilizing effect of stock enhancement to emerge. The first assumption is that released fish compete with wild fish. This type of competition is very likely to occur in the study streams. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that this species compete with other species of stream fish. In particular, masu salmon raised in hatchery environments are more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild. Second, our theory predicts the emergence of destabilizing effect when the number of fish released is externally controlled. This is exactly what was observed in our study system. In the protected watersheds, there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). Instead, these released fish are likely controlled by socioeconomic factors, such as budget allocation to hatchery programs in this region. As such, released fish are probably "excessive" given the density-dependent regulation of natural communities, imposing intense resource competition that would otherwise not exist.

Like any non-experimental studies, we cannot exclude the possibility that our results were driven by spurious correlations. However, this is very unlikely because we have controlled potential confounding factors, including climatic and land use variables. In addition, our results were in close agreement with theoretical predictions, which are free from any confounding factors. Hence, our empirical results are sufficiently reliable.

Although captive breeding programs pay close attention to the fate of captive-bred individuals in the wild, it is rare to consider the ecosystem's capacity to support biodiversity. We anticipate that the destabilizing effect of stock enhancement is pervasive across taxa because negative density dependence is a general regulating mechanism of biodiversity. In particular, resource competition may be intensified in the Anthropocene because human influences have caused substantial loss and degradation of natural habitats. Therefore, one important implication of this study is that priorities should be given to the preservation and restoration of natural habitats with a broader focus on biodiversity. For example, recent research underscores the importance of conserving a diverse habitat portfolio that may buffer the impacts of rapid environmental changes. Without a comprehensive framework that appreciates the ecological integrity of self-sustaining communities, stock enhancement will never be effective but impairs biodiversity.
